Manuscript title here

### 2021-08-05 docs/ms/ms.RMD
updated manuscript and workflow.

source(here::here("datascript.R"))

@@

```
<!--- this is a note in Rmarkdown --->

@@

Trawl sampling was conducted as part of the Louisiana Universities Marine Consortium's (LUMCON) educational outreach program (https://www.lumcon.edu/education-and-outreach). Individual trawls were conducted using *trawl apparatus description...* and occasionally multiple

@@

Species names were checked for typos and confirmed using the *gni_parse* function from the *taxize* package [@chamberlain2020] and further standardized to consistent taxonomic resolution for entire data series (Supplemental materials).

mobr package [@mcglinn2019]

Trawls with >20 individuals counted (This should be reconsidered when we aggregate at the annual scale)
Bootstrapping methods were used by resampling with replacement `r nperm` times. From these permutations the 2.5% and 97.5% percentiles of yearly statistics were taken to estimate 95% percentile intervals. With this procedure we estimate:
- *S~asymp~* for each year using the Chao1 estimator [@chao1984][@chao1987] a metric which is highly correlated with richness, *S* [@mcgill2011]
- because of differences in effort, we estimated effort-controlled rarefied species richness, *S~n~*, which provides the expected number of species given a defined number of *n* individuals sampled [@gotelli2001]. *S~n~* estimates for yearly aggregated data, *n* was calculated by the minimum samples within a year multipled by `r n_year`. 
- We used the hillR package [@li2018] to calculate the diversity order 2, which corresponds to the effective number of dominant (highly abundant) species.

@@

Over the sampling period from `r min(TB_trawl_sampling$year, na.rm = TRUE)` to `r max(TB_trawl_sampling$year, na.rm = TRUE)`, `r length(unique(TB_trawl_data$Date))` trawls-days were performed yielding `r format(signif(sum(TB_trawl_data$Abundance, na.rm = TRUE),3), scientific = TRUE)` individuals. Among years, total effort generally increased, both in number of trawl-days and individuals counted, from a minimum effort of `r min(TB_trawl_sampling$trawl_days)` (`r TB_trawl_sampling %>% slice_min(trawl_days) %>% dplyr::select(year) %>% unlist`) to `r max(TB_trawl_sampling$trawl_days)` (`r TB_trawl_sampling %>% slice_max(trawl_days) %>% dplyr::select(year) %>% unlist`) days and from `r min(TB_trawl_sampling$individuals, na.rm = TRUE)` (`r TB_trawl_sampling %>% slice_min(individuals) %>% dplyr::select(year) %>% unlist`) to `r format(signif(max(TB_trawl_sampling$individuals, na.rm = TRUE),3), scientific = TRUE)` (`r TB_trawl_sampling %>% slice_max(individuals) %>% dplyr::select(year) %>% unlist`) individuals. In total, `r length(unique(TB_trawl_data$Common_name))` unique species were identified and new species were consistently observed throughout the sampling period (Figure 1). Within a calendar year, we estimated the asymptotic species richness (*S~asymp~*) which was relatively stable for the first five years (2007 -- 2012), reaching a minimum value of `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(median) %>% unlist %>% as.numeric %>% round(0)` (`r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(lower) %>% unlist %>% as.numeric %>% round(0)` -- `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(upper) %>% unlist %>% as.numeric %>% round(0)`) in `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(group)`. Estimated *S~asymp~* increased in 2013 to a series maximum of `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(median) %>% unlist %>% as.numeric %>% round(0)` (`r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(lower) %>% unlist %>% as.numeric %>% round(0)` -- `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(upper) %>% unlist %>% as.numeric %>% round(0)`) in `r group_stats %>% data.frame %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(group)`. However, given the large differences in effort among years, variability in *S~asymp~* may be partly attributable to sampling effects among years. Therefore, we also calculated rarefied richness at `r S_n` individuals, *S~`r S_n`~*,  for each year. Here, effort-controlled species richness, *S~`r S_n`~*, diverged qualitatively from *S~asymp~* and was generally constant, ranging from `r group_stats %>% dplyr::filter(effort > 100) %>% slice_min(median) %>% dplyr::select(median) %>% round(1)` (`r group_stats %>% dplyr::filter(effort > 100) %>% slice_min(median) %>% dplyr::select(lower) %>% round(1)` -- `r group_stats %>% dplyr::filter(effort > 100) %>% slice_min(median) %>% dplyr::select(upper) %>% round(1)`) to `r group_stats %>% dplyr::filter(effort > 100) %>% slice_max(median) %>% dplyr::select(median) %>% round(1)` (`r group_stats %>% dplyr::filter(effort > 100) %>% slice_max(median) %>% dplyr::select(lower) %>% round(1)` -- `r group_stats %>% dplyr::filter(effort > 100) %>% slice_max(median) %>% dplyr::select(upper) %>% round(1)`) (Figure 1).

@@

# partition beta diversity into nestedness and turnover #
# uses `betapart` package #
TB_time_df <- TB_time_accum %>%
  dplyr::select(Date, year, richness) %>%
  mutate(data_type = 'continuous',
         r_lower = NA_real_,
         r_upper = NA_real_) %>%
  bind_rows(TB_annual_accum %>% 
              mutate(data_type = "ann_raw",
                     r_lower = NA_real_,
                     r_upper = NA_real_)) %>%
  bind_rows(group_stats %>% 
              dplyr::filter(effort > 100) %>%
              dplyr::select(year ='group', richness = 'median', r_lower = 'lower', r_upper = 'upper') %>%
              dplyr::mutate(year = as.numeric(as.character(year)),
                            Date = as.Date(paste(year,"01","01", sep = "-"), format = "%Y-%m-%d"),
                            data_type = "rarefied")) %>%
  bind_rows(group_stats %>% 
              dplyr::filter(index == 'S_asymp') %>%
              dplyr::select(year ='group', richness = 'median', r_lower = 'lower', r_upper = 'upper') %>%
              dplyr::mutate(year = as.numeric(as.character(year)),
                            Date = as.Date(paste(year,"01","01", sep = "-"), format = "%Y-%m-%d"),
                            data_type = "ann_asymp")) %>%
  dplyr::mutate(data_type = factor(data_type, levels = c("continuous", "rarefied", "ann_raw", "ann_asymp"))) %>%
  dplyr::filter(data_type != "ann_raw")

rich_time <-
  ggplot(TB_time_df, aes(x = Date, y = richness, group = data_type)) +
  geom_ribbon(data = TB_time_df %>%
                dplyr::filter(data_type == "ann_asymp"),
              aes(x = Date, ymin = r_lower, ymax= r_upper),
              color = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')[3],
              fill = viridis(n = 3, alpha = c(1,1,0.1), option = 'viridis')[3],
              inherit.aes = FALSE)+
  geom_ribbon(data = TB_time_df %>%
                dplyr::filter(data_type == "rarefied"),
              aes(x = Date, ymin = r_lower, ymax= r_upper),
              color = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')[2],
              fill = viridis(n = 3, alpha = c(1,0.2,1), option = 'viridis')[2],
              inherit.aes = FALSE)+
  geom_segment(data = TB_time_df %>% 
                 dplyr::filter(data_type %ni% 'continuous') %>%
                 pivot_wider(Date, names_from = data_type, values_from = richness),
               aes(x = Date, xend = Date, y = ann_asymp, yend = rarefied ), 
               color = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')[3],
               inherit.aes = F) +
  geom_point(data = TB_time_df, aes(colour = data_type),size = 1.5) +
  geom_line(data = TB_time_df %>% dplyr::filter(!grepl('ann_asymp', data_type)), aes(colour = data_type),size = 1.2) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(name = "Species Richness")+
  scale_colour_manual(labels = c("continuous", expression(italic("S"[900])),expression(italic("S"["asymp"]))), values = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')) +
  # scale_alpha_manual(values = c(1,1,0.3))+
  theme(legend.position = c(0, 1), legend.justification = c(0, 1), legend.background = element_rect(fill = "transparent", colour = NA),
        legend.title = element_blank(), legend.text.align = 0,
        # axis.title.x = element_blank(), axis.text.x = element_blank(),
        legend.spacing.y = unit(-0.1, 'cm'))

@@

Considering the consistent accumulation of new species throughout the sampling series and the relatively stable effort-controlled richness, we

```{r figure 2, echo=FALSE, warning=FALSE, message=FALSE}

TB_trawl_taxasite_yr <- TB_trawl_taxasite %>%
  dplyr::select(-julian_date,-Date,-month,-biweekly) %>%
  group_by(year) %>%
  dplyr::mutate(year = as.Date(paste(year,"01","01", sep = "-"), format = "%Y-%m-%d")) %>%
  dplyr::summarise(across(where(is.numeric), ~sum(.x, na.rm = TRUE)))

# TB_trawl_yr_qD = TB_trawl_taxasite_yr %>%
#   split(nrow(.)) %>%
#   purrr::map2(., c(0,1,2), ~hillR::hill_taxa(.x %>% dplyr::select(-year), q = .y)) %>% 
#   setNames(., nm = c("q0","q1","q2")) %>%
#   bind_cols() %>%
#   dplyr::mutate(year = TB_trawl_taxasite_yr$year)

# TB_trawl_yr_qD %>%
# # group_stats %>%
# #   dplyr::filter(index == "S_PIE") %>%
# #   dplyr::mutate(group = as.Date(group, format = "%Y")) %>%
#   ggplot()+
#   geom_point(aes(x = year, y = q2))+
#   geom_path(aes(x = year, y = q2))+
#   # geom_point(aes(x = year, y = q1))+
#   # geom_path(aes(x = year, y = q1))+
#   # geom_ribbon(aes(x = group, ymin = lower, ymax = upper), alpha = 0.3)+
#   scale_color_manual(values = viridis(n =2, option = 'viridis' ))+
#   scale_fill_manual(values = viridis(n = 2, option = 'viridis'))+
#   scale_y_continuous(limits = c(0,10))
q2_boots = boot_sample_groups(TB_trawl_taxasite %>% dplyr::select(-julian_date,-Date,-month,-biweekly), 
                   group_var = 'year',
                   qD = 2,
                   n_perm = nperm)

q1_boots = boot_sample_groups(TB_trawl_taxasite %>% dplyr::select(-julian_date,-Date,-month,-biweekly), 
                   group_var = 'year',
                   qD = 1,
                   n_perm = nperm)

group_stats %>%
  dplyr::filter(index == "pct_rare") %>%
  dplyr::mutate(group = as.Date(group, format = "%Y")) %>%
  ggplot()+
  geom_point(aes(x =group, y = median))+
  geom_path(aes(x = group, y = median))+
  geom_ribbon(aes(x = group, ymin = lower, ymax = upper), alpha = 0.3)+
  scale_color_manual(values = viridis(n =1, option = 'viridis' ))+
  scale_fill_manual(values = viridis(n = 1, option = 'viridis'))#+
  # scale_y_continuous(limits = c(0,10))

>>>>

source(here::here("datascript.R"))
source(here::here("analyses/trawl/R/Lorenz.R"))

@@

```
<!--- this is a note in Rmarkdown --->
\newpage

@@

Trawl sampling was conducted as part of the Louisiana Universities Marine Consortium's (LUMCON) educational outreach program (https://www.lumcon.edu/education-and-outreach). Individual trawls were conducted using *trawl apparatus description...* for 10 - 20 min. When multiple trawls were conducted on a single day, species counts were summed across all trawls and

@@

Species names were checked for typos and confirmed using the *gni_parse* function from the *taxize* package [@chamberlain2020] and further standardized to consistent taxonomic resolution for entire data series (Supplemental materials). We measured multiple aspects of species richness, first to track the uncorrected accumulation of species over time we filtered species from the data set after first occurrence. We then aggreagated all trawls within a calendar year and calculated two measures of species richness, accounting for differences in sampling effort among years. The first measure was asymptotic species richness, *S~asymp~*, using the Chao1 estimator [@chao1984][@chao1987]. This metric is highly correlated with measured richness, *S* [@mcgill2011], but attempts to extrapolate the total species richness based on incomplete sampling. Second, we attempted to account for differences in effort among years by estimating the effort-controlled rarefied species richness, *S~n~*, which provides the expected number of species given a defined number of *n* individuals sampled [@gotelli2001]. *S~n~* estimates for yearly aggregated data, *n* was calculated by the minimum samples within a year multiplied by `r n_year`. We calculated *S~asymp~* and *S~n~* using the mobr package [@mcglinn2019][@mcglinn2021a].

Beyond actual and expected numbers of species, we calculated additional measures of biodiversity to capture community change through time. To quantify changes in the distribution of abundant and rare species, we calculated specific orders of diversity, the percentage of rare species, and evenness in species abundances on annually aggregated species abundances. An order of diversity, *^q^D*, where *q* represents a non-negative integer, captures the "effective numbers of species" based on varying degrees of rarity of the species in a community [@hill1973]. We used the the *hill_taxa* function in the hillR package [@li2018] to estimate the effective species for diversity order 2 (i.e., *^2^D*). *^2^D* is equivalent to the reciprocal of Simpson's diversity [@simpson1949] (i.e., 1/Simpson's index) which emphasizes the importance of dominant (highly abundant) species. At the other end of the abundance spectrum, we calculated the percentage of rare species in the community as the number of species with abundances below a threshold of $0.05*n$, divided by the total number of species x 100. Here, *n* is the number of individuals in the sample and the value of 0.05 was chosen because it has been shown to capture rarity well and is generally uncorrelated with other metrics of biodiversity. Lastly, to integrate changes in abundance of dominant and rare species within the community, we calculated the evenness of species abundances with the Gini coefficient [@gini1921], normalized for differences in species richness among years, $G^*$ [@solomon1975; @chao2019], :

$$ G^* = (2 \sum_{i = 1}^S ip_i -2)/(S-1)$$

where, in a community of *S* species, species' relative abundances, $p_{i}$, are ordered such that $p_{1} \geq p_{2} \geq ...p_{S}$ for all species of the community. For all metrics, variability was calculated with bootstrapping methods by resampling with replacement `r nperm` times. From these permutations the 2.5% and 97.5% percentiles of yearly statistics were taken to estimate 95% percentile intervals.

@@

Over the sampling period from `r min(TB_trawl_sampling$year, na.rm = TRUE)` to `r max(TB_trawl_sampling$year, na.rm = TRUE)`, `r length(unique(TB_trawl_data$Date))` trawls-days were performed yielding `r format(signif(sum(TB_trawl_data$Abundance, na.rm = TRUE),3), scientific = TRUE)` individuals. Among years, total effort generally increased, both in number of trawl-days and individuals counted, from a minimum effort of `r min(TB_trawl_sampling$trawl_days)` (`r TB_trawl_sampling %>% slice_min(trawl_days) %>% dplyr::select(year) %>% unlist`) to `r max(TB_trawl_sampling$trawl_days)` (`r TB_trawl_sampling %>% slice_max(trawl_days) %>% dplyr::select(year) %>% unlist`) days and from `r min(TB_trawl_sampling$individuals, na.rm = TRUE)` (`r TB_trawl_sampling %>% slice_min(individuals) %>% dplyr::select(year) %>% unlist`) to `r format(signif(max(TB_trawl_sampling$individuals, na.rm = TRUE),3), scientific = TRUE)` (`r TB_trawl_sampling %>% slice_max(individuals) %>% dplyr::select(year) %>% unlist`) individuals. In total, `r length(unique(TB_trawl_data$Common_name))` unique species were identified and new species were consistently observed throughout the sampling period (Figure 1). Within a calendar year, we estimated the asymptotic species richness (*S~asymp~*) which was relatively stable for the first five years (2007 -- 2012), reaching a minimum value of `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(median) %>% unlist %>% as.numeric %>% round(0)` (`r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(lower) %>% unlist %>% as.numeric %>% round(0)` -- `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(upper) %>% unlist %>% as.numeric %>% round(0)`) in `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(group)`. Estimated *S~asymp~* increased in 2013 to a series maximum of `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(median) %>% unlist %>% as.numeric %>% round(0)` (`r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(lower) %>% unlist %>% as.numeric %>% round(0)` -- `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(upper) %>% unlist %>% as.numeric %>% round(0)`) in `r group_stats %>% data.frame %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(group)`. However, given the large differences in effort among years, variability in *S~asymp~* may be partly attributable to sampling effects among years. Therefore, we also calculated rarefied richness at `r S_n` individuals, *S~`r S_n`~*,  for each year. Here, effort-controlled species richness, *S~`r S_n`~*, diverged qualitatively from *S~asymp~* and was generally constant with an apparent slight increase in more recent sampling years. Overall, *S~`r S_n`~* ranged from `r group_stats %>% dplyr::filter(effort > 100) %>% slice_min(median) %>% dplyr::select(median) %>% round(1)` (`r group_stats %>% dplyr::filter(effort > 100) %>% slice_min(median) %>% dplyr::select(lower) %>% round(1)` -- `r group_stats %>% dplyr::filter(effort > 100) %>% slice_min(median) %>% dplyr::select(upper) %>% round(1)`) to `r group_stats %>% dplyr::filter(effort > 100) %>% slice_max(median) %>% dplyr::select(median) %>% round(1)` (`r group_stats %>% dplyr::filter(effort > 100) %>% slice_max(median) %>% dplyr::select(lower) %>% round(1)` -- `r group_stats %>% dplyr::filter(effort > 100) %>% slice_max(median) %>% dplyr::select(upper) %>% round(1)`) (Figure 1).

@@

### create timeseries figure
TB_time_df <- TB_time_accum %>%
  dplyr::select(Date, year, richness) %>%
  mutate(data_type = 'continuous',
         r_lower = NA_real_,
         r_upper = NA_real_) %>%
  bind_rows(TB_annual_accum %>% 
              mutate(data_type = "ann_raw",
                     r_lower = NA_real_,
                     r_upper = NA_real_)) %>%
  bind_rows(group_stats %>% 
              dplyr::filter(effort > 100) %>%
              dplyr::select(year ='group', richness = 'median', r_lower = 'lower', r_upper = 'upper') %>%
              dplyr::mutate(year = as.numeric(as.character(year)),
                            Date = as.Date(paste(year,"01","01", sep = "-"), format = "%Y-%m-%d"),
                            data_type = "rarefied")) %>%
  bind_rows(group_stats %>% 
              dplyr::filter(index == 'S_asymp') %>%
              dplyr::select(year ='group', richness = 'median', r_lower = 'lower', r_upper = 'upper') %>%
              dplyr::mutate(year = as.numeric(as.character(year)),
                            Date = as.Date(paste(year,"01","01", sep = "-"), format = "%Y-%m-%d"),
                            data_type = "ann_asymp")) %>%
  dplyr::mutate(data_type = factor(data_type, levels = c("continuous", "rarefied", "ann_raw", "ann_asymp"))) %>%
  dplyr::filter(data_type != "ann_raw")

S_n_lab = paste0("S[",S_n,"]")

rich_time <-
  ggplot(TB_time_df, aes(x = Date, y = richness, group = data_type)) +
  geom_ribbon(data = TB_time_df %>%
                dplyr::filter(data_type == "ann_asymp"),
              aes(x = Date, ymin = r_lower, ymax= r_upper),
              color = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')[3],
              fill = viridis(n = 3, alpha = c(1,1,0.1), option = 'viridis')[3],
              inherit.aes = FALSE)+
  geom_ribbon(data = TB_time_df %>%
                dplyr::filter(data_type == "rarefied"),
              aes(x = Date, ymin = r_lower, ymax= r_upper),
              color = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')[2],
              fill = viridis(n = 3, alpha = c(1,0.2,1), option = 'viridis')[2],
              inherit.aes = FALSE)+
  geom_segment(data = TB_time_df %>% 
                 dplyr::filter(data_type %ni% 'continuous') %>%
                 pivot_wider(Date, names_from = data_type, values_from = richness),
               aes(x = Date, xend = Date, y = ann_asymp, yend = rarefied ), 
               color = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')[3],
               inherit.aes = F) +
  geom_point(data = TB_time_df, aes(colour = data_type),size = 1.5) +
  geom_line(data = TB_time_df %>% dplyr::filter(!grepl('ann_asymp', data_type)), aes(colour = data_type),size = 1.2) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(name = "Species Richness")+
  scale_colour_manual(labels = c("continuous", bquote(italic("S"[.(S_n[1])])),expression(italic("S"["asymp"]))), values = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')) +
  # scale_alpha_manual(values = c(1,1,0.3))+
  theme(legend.position = c(0, 1), legend.justification = c(0, 1), legend.background = element_rect(fill = "transparent", colour = NA),
        legend.title = element_blank(), legend.text.align = 0,
        # axis.title.x = element_blank(), axis.text.x = element_blank(),
        legend.spacing.y = unit(-0.1, 'cm'))+
  scale_x_date(date_breaks = 'year', limits = c(as.Date("2007-01-01"),as.Date("2019-06-01")),
               date_labels = "%Y")

@@

```{r turnover, echo = FALSE, warning=FALSE, message=FALSE}

## quantify loss and appearance of species in the timeseries
TB_trawl_data %>%
  # exclude 2020 & 2021
  dplyr::filter(year %ni% c("2020","2021")) %>%
  group_by(year, Common_name) %>%
  dplyr::summarise(Abundance = sum(Abundance, na.rm = TRUE)) %>%
  group_by(year) %>%
  pivot_wider( names_from = 'Common_name', values_from = 'Abundance', values_fill = 0) %>%
  ungroup %>%
  # dplyr::mutate(time.var = 1:n()) %>%
  pivot_longer(c(-year), names_to = 'Common_name', values_to = 'Abundance')-> TB_trawl_yr

trawl_yr_turnover = purrr::map(c('appearance','disappearance','total'),
                                     ~codyn::turnover(df = TB_trawl_yr,
                                                      time.var = 'year',
                                                      species.var = 'Common_name',
                                                      abundance.var = 'Abundance',
                                                      metric = .x) %>%
                                       data.frame %>%
                                       dplyr::mutate(Date = as.Date(paste0(unique(TB_trawl_yr$year)[-1],"-01-01"), format = "%Y-%m-%d"),
                                                     variable = .x) %>%
                                       dplyr::rename(value = !!names(.[1])) %>%
                                       bind_rows(data.frame(value = NA_real_, time.var = NA_real_, Date = as.Date("2008-01-01"), variable =.x))) %>%
                                     bind_rows
gains = trawl_yr_turnover %>%
  dplyr::filter(variable == "appearance")

losses = trawl_yr_turnover %>%
  dplyr::filter(variable == "disappearance")

total = trawl_yr_turnover %>%
  dplyr::filter(variable == "total")

Consistent with the pattern of continued accumulation of new species throughout the sampling series and the relatively stable effort-controlled richness, species gains and losses generally balanced (Figure 2 ). Total turnover averaged `r round(mean(total$value, na.rm = TRUE),2)` and decreased with time from a peak turnover of `r round(max(total$value, na.rm = TRUE),2)` in `r total %>% slice_max(value) %>% dplyr::select(year)` to `r total %>% dplyr::filter(year == '2019') %>% dplyr::select(value) %>% round(2)` in the final year of record, 2019 (Figure 2). The contribution of species gains to total turnover averaged `r round(mean(gains$value, na.rm = TRUE),2)` and ranged from `r round(max(gains$value, na.rm = TRUE),2)` to `r round(min(gains$value, na.rm = TRUE),2)`, whereas, species losses average `r round(mean(losses$value, na.rm = TRUE),2)` and ranged from `r round(max(losses$value, na.rm = TRUE),2)` to `r round(min(losses$value, na.rm = TRUE),2)`.

```{r turnover plot ,echo = FALSE, warning=FALSE, message=FALSE, fig.cap = "Figure 2. Total species "}
trawl_yr_turnover %>%
  ggplot()+
  geom_line(aes(x = Date, y = value, group = variable, color = variable), size = 1.1)+
  scale_color_manual(values = viridis(n = 3, option = 'viridis'))+
  scale_y_continuous(name = 'Species turnover',limits = c(0,0.7), expand = c(0.01,0.01))+
  scale_x_date(date_breaks = 'year', limits = c(as.Date("2007-01-01"),as.Date("2019-06-01")),
               date_labels = "%Y")+
  theme(legend.position = c(0,0),
        legend.justification = c(0,0),
        legend.background = element_rect(fill = 'transparent', color = NA),
        legend.title = element_blank(),
        axis.title.x = element_blank())
```

```{r annual qD, echo=FALSE, warning=FALSE, message=FALSE}
set.seed(123)
q2_boots = boot_sample_groups(TB_trawl_taxasite %>% dplyr::select(-julian_date,-Date,-month,-biweekly), 
                   group_var = 'year',
                   qD = 2,
                   n_perm = nperm)
set.seed(123)
q1_boots = boot_sample_groups(TB_trawl_taxasite %>% dplyr::select(-julian_date,-Date,-month,-biweekly), 
                              group_var = 'year',
                              qD = 1,
                              n_perm = nperm)

qD_df = q2_boots %>% 
  # summarise q2 by year
  group_by(year) %>%
  dplyr::summarise(across(q2, list(mean = ~mean(.x, na.rm = TRUE),
                                   median = ~quantile(.x, 0.5, na.rm = TRUE),
                                   quant2.5= ~quantile(.x, 0.025, na.rm = TRUE),
                                   quant25= ~quantile(.x, 0.25, na.rm = TRUE),
                                   quant75= ~quantile(.x, 0.75, na.rm = TRUE),
                                   quant97.5= ~quantile(.x, 0.975, na.rm = TRUE)), .names = "{.fn}")) %>%
  dplyr::mutate(Date = as.Date(paste(year,"01","01", sep = "-"), format = "%Y-%m-%d"),
                q_order = "q2") %>%
  bind_rows(
    q1_boots %>%
      # summarise q1 by year
      group_by(year) %>%
      dplyr::summarise(across(q1, list(mean = ~mean(.x, na.rm = TRUE),
                                       median = ~quantile(.x, 0.5, na.rm = TRUE),
                                       quant2.5= ~quantile(.x, 0.025, na.rm = TRUE),
                                       quant25= ~quantile(.x, 0.25, na.rm = TRUE),
                                       quant75= ~quantile(.x, 0.75, na.rm = TRUE),
                                       quant97.5= ~quantile(.x, 0.975, na.rm = TRUE)), .names = "{.fn}")) %>%
      dplyr::mutate(Date = as.Date(paste(year,"01","01", sep = "-"), format = "%Y-%m-%d"),
                    q_order = "q1")
   )

annual_evenness_boots = boot_gini_groups(TB_trawl_taxasite %>% dplyr::select(-julian_date,-Date,-month,-biweekly), 
                                group_var = 'year',
                                n_perm = nperm)

annual_evenness_df = annual_evenness_boots %>%
  dplyr::select(-boot_id) %>%
  pivot_longer(-year, names_to = 'measure', values_to = 'value') %>%
  group_by(year,measure) %>%
  dplyr::summarise(across(value, list(mean = ~mean(.x, na.rm = TRUE),
                                      median = ~quantile(.x, 0.5, na.rm = TRUE),
                                      quant2.5= ~quantile(.x, 0.025, na.rm = TRUE),
                                      quant25= ~quantile(.x, 0.25, na.rm = TRUE),
                                      quant75= ~quantile(.x, 0.75, na.rm = TRUE),
                                      quant97.5= ~quantile(.x, 0.975, na.rm = TRUE)), .names = "{.fn}")) %>%
  dplyr::mutate(Date = as.Date(paste(year,"01","01", sep = "-"), format = "%Y-%m-%d")) %>%
  ungroup

group_stats %>%
  dplyr::filter(index == "pct_rare") -> rarity

Species turnover was likley confined to species with moderate to low relative abundances. The effective number of highly abundant species, *^2^D*, was variable but stable through time with calculated values *ca.* `r qD_df %>% dplyr::filter(q_order == 'q2') %>% dplyr::summarise(across(median, ~median(.x ,na.rm = TRUE))) %>% round(2)` $\pm$ `r qD_df %>% dplyr::filter(q_order == 'q2') %>% dplyr::summarise(across(median, ~sd(.x, na.rm = TRUE))) %>% round(1)` (mean $\pm$ 1 SD) (range: `r qD_df %>% dplyr::filter(q_order == 'q2')%>% slice_min(median) %>% dplyr::select(median) %>% round(2)` to  `r qD_df %>% dplyr::filter(q_order == 'q2') %>% slice_max(median) %>% dplyr::select(median) %>% round(1)`; Figure 3 __*supplemental?*__). In contrast, the percentage of rare species increased through time from `r rarity %>% slice_min(group) %>% dplyr::select(median) %>% round(2)` (`r rarity %>% slice_min(group) %>% dplyr::select(lower) %>% round(2)` -- `r rarity %>% slice_min(group) %>% dplyr::select(upper) %>% round(2)`, 95% percentile interval [PI]) to `r rarity %>% slice_max(group) %>% dplyr::select(median) %>% round(2)` `r rarity %>% slice_max(group) %>% dplyr::select(lower) %>% round(2)` -- `r rarity %>% slice_max(group) %>% dplyr::select(upper) %>% round(2)`, 95% PI; Figure 3A). Additionally, the normalized evenness in species relative abundances showed a consistent decline through time from a median value of `r annual_evenness_df %>% dplyr::filter(measure == 'gini_norm') %>% slice_min(Date) %>% dplyr::select(median) %>% round(2)` in `r min(annual_evenness_df$year,na.rm = TRUE)` (`r annual_evenness_df %>% dplyr::filter(measure == 'gini_norm') %>% slice_min(Date) %>% dplyr::select(quant2.5) %>% round(2)` -- `r annual_evenness_df %>% dplyr::filter(measure == 'gini_norm') %>% slice_min(Date) %>% dplyr::select(quant97.5) %>% round(2)`, 95% PI) to `r annual_evenness_df %>% dplyr::filter(measure=='gini_norm') %>% slice_max(Date) %>% dplyr::select(median) %>% round(2)` (`r annual_evenness_df %>% dplyr::filter(measure == 'gini_norm') %>% slice_max(Date) %>% dplyr::select(quant2.5) %>% round(2)` -- `r annual_evenness_df %>% dplyr::filter(measure == 'gini_norm') %>% slice_max(Date) %>% dplyr::select(quant97.5) %>% round(2)`, 95% PI) in `r max(annual_evenness_df$year, na.rm = TRUE)`.

```{r diversity2 plot, echo = FALSE, warning=FALSE, message=FALSE, fig.cap="Figure 3. Temporal trends in order of diversity q = 2, ^2^D highlight relative stability in the number of dominant species in the community through time. ^2^D is equivalent to 1/Simpson's diversity index and captures the effective number of highly abundant species in the community."}
qD_df %>%
  dplyr::filter(q_order == "q2") %>%
  ggplot()+
  geom_ribbon(aes(x = Date, ymin = quant2.5, ymax = quant97.5, group = q_order, fill = q_order), alpha = 0.5)+
  geom_line(aes(x = Date, y = median, group = q_order, color = q_order), size = 1.2)+
  geom_point(aes(x = Date, y= mean, group = q_order), shape = 6)+
  scale_y_continuous(name = expression("Effective species ("~italic(''^2*D)~")"),limits = c(0,11))+
  scale_color_manual(labels = c("q2"), values = viridis(n = 1, option = 'viridis'))+
  scale_fill_manual(labels = c("q2"),values = viridis(n = 1, alpha = 0.5, option = 'viridis'))+
  theme(legend.title = element_blank(),
        legend.position = "none",
        legend.justification = c(0,1))

```{r evenness, echo = FALSE, warning=FALSE, message=FALSE, fig.cap= "Figure 4." }

annual_evenness_df %>%
  dplyr::filter(measure == 'gini_raw') %>%
  ggplot()+
  geom_ribbon(aes(x = Date, ymin = quant2.5, ymax = quant97.5), alpha = 0.3) +
  geom_line(aes(x = Date, y = median), size = 1.1) +
  geom_point(aes(x = Date, y = median), shape = 22, size = 3)+
  scale_y_continuous(name = "Normalized Gini coefficient", limits = c(0,NA), expand = c(0.01,0.01))+
  scale_x_date(date_breaks = 'year', limits = c(as.Date("2007-01-01"),as.Date("2019-06-01")),
               date_labels = "%Y")+
  theme(axis.title.x = element_blank())-> evenness_plot

group_stats %>%
  dplyr::filter(index == "pct_rare") %>%
  dplyr::mutate(group = as.Date(group, format = "%Y")) %>%
  ggplot()+
  geom_point(aes(x =group, y = median), size = 3, shape = 22)+
  geom_path(aes(x = group, y = median), size = 1.1)+
  geom_ribbon(aes(x = group, ymin = lower, ymax = upper), alpha = 0.3)+
  scale_y_continuous(name = "Percent rare species (%)", limits = c(70,100), expand = c(0.01,0.01))+
  scale_color_manual(values = viridis(n =1, option = 'viridis' ))+
  scale_fill_manual(values = viridis(n = 1, option = 'viridis'))+
  scale_x_date(date_breaks = 'year', limits = c(as.Date("2007-01-01"),as.Date("2019-06-01")),
               date_labels = "%Y")+
  theme(axis.title.x = element_blank())-> rarity_plot

gridExtra::grid.arrange(rarity_plot, evenness_plot, ncol = 1)
